---
title: "Capstone_Project_Analysis_Report"
author: "Jumana Alaarajee"
date: "2025-03-10"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

## Introduction: 
In this analysis we explore social media engagement across various social media platforms such as Facebook, Instagram,
TikTok and many others. Engagement is defined as the combination of user
activity which includes likes, posts, follows and time spent on the
application. The objective of this analysis is to determine which social
media platform has the best engagement and assess any correlation
between user interaction (likes, follows, posts) and time spent.

## Data Upload: 

The dataset used for this analysis can be obtained through Kaggle,
<https://www.kaggle.com/datasets/bhadramohit/social-media-usage-datasetapplications>.
This dataset takes information from 1000 random users across all social
media platforms and provides information regarding time spend on the
app, likes, follows, and posts.

```{r}

data <-read.csv("social_media_usage.csv")
head(data)

```

## Engagement Calculation:

We define engagement as a ratio of user activity (posts, likes, and
follows) to the time spent on the platform. The formula for calculating
engagement is:

Engagement = (Posts Per Day + Likes Per Day + Follows Per Day) / Daily
Minutes Spent

This provides us with a normalized measure of user activity. We create a
new column, engagement, to store this calculated value for each user.

```{r, echo=FALSE}
# calculate engagement
data$engagement <- (data$Posts_Per_Day+data$Likes_Per_Day+data$Follows_Per_Day)/data$Daily_Minutes_Spent 
# check first couple of rows of dataset, can be uncommented based on readers preference
#head(data)
```

## Engagement Barplot Visualization:

Next, we visualize the average engagement per social media platform. The
bar plot below shows the average engagement across different platforms,
helping us identify which platform has the highest average user
engagement.

```{r, echo=FALSE}
# mapped average engagement on barplot 
library(ggplot2)
ggplot(data, aes(x=App, y=engagement, color=App))+geom_bar(stat = "summary", fun = "mean")+ labs(title = "Average Engagment per Social Media Platform", x="App", y="Avg. Engagement")
```

## Correlation Analysis:

To further understand the relationships between user behavior and
engagement, we calculate the correlations between various factors: posts
per day, likes per day, follows per day, and daily minutes spent. These
correlations help us understand the strength of the relationship between
different user interactions and time spent on the platform.

```{r}
cor(data$Posts_Per_Day,data$engagement)
cor(data$Likes_Per_Day,data$engagement)
cor(data$Follows_Per_Day,data$engagement)
cor(data$Daily_Minutes_Spent,data$engagement)
```

## Correlation between other Attributes

```{r}
cor(data$Posts_Per_Day,data$Likes_Per_Day)
cor(data$Posts_Per_Day,data$Follows_Per_Day)
cor(data$Posts_Per_Day,data$Daily_Minutes_Spent)

cor(data$Daily_Minutes_Spent, data$Likes_Per_Day)
cor(data$Daily_Minutes_Spent, data$Follows_Per_Day)

cor(data$Likes_Per_Day, data$Follows_Per_Day)
```

## Analysis Findings:

Engagement and User Interactions: There is no significant correlation
between engagement and individual user interactions such as posts,
likes, or follows. This suggests that engagement may not be entirely
dependent on these factors and can be influenced by other factors not
mentioned, such as user preferences or content quality.

Time Spent vs User Interactions: Similarly, the correlation between time
spent and user interactions (likes, follows, posts) is also weak. This
indicates that users who spend more time on the platform do not
necessarily engage more in terms of likes, follows, or posts.

Overall Trends: The correlations between the different user interactions
(posts, likes, follows) are relatively weak, indicating that these
factors may be independent of each other.

## Conclusion:

Average Engagement: Different platforms have varying levels of average
engagement, with certain platforms showing higher levels of user
interaction compared to others.We find that Instagram and TikTok exhibit
the two highest average user engagement. This could be explained by the
fact that these platforms are particularly popular among younger
generations, who are typically more active on social media.
Understanding these trends can help businesses or content creators
identify platforms with the highest user engagement, aiding in more
effective outreach and audience targeting.

Correlations: The lack of significant correlations between engagement
and user behavior metrics (posts, likes, follows) implies that these
factors do not always lead to higher engagement. Additionally, the weak
correlations between time spent and user interactions suggest that other
factors might be influencing engagement.
